Entropy-SGD: Biasing Gradient Descent Into Wide Valleys

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
minor, adding testing code for char-rnn

In order to test the efficacy of $\entropysgd$ on recurrent architectures, we train an LSTM network on the Penn Tree Bank (PTB) dataset for word-level text prediction. This dataset contains about one million words divided into a training set of about $930,000$ words, a validation set of $74,000$ words and $82,000$ words with a vocabulary of size $10,000$. As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}. Our network called $\ptblstm$ consists of two layers with $1500$ hidden units each with a dropout of probability $0.65$; this is a large network with about $66$ million weights. It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch. \pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which matches that of the original authors.

@@

Next,

\todo{Let us note that tuning the momentum hyper-parameter was crucial to getting good results on RNNs. While the SGD baseline on $\ptblstm$ does not use momentum and in fact does not train well with momentum, we set it to $0.5$ for $\entropysgd$. On the other hand, the baseline for $\charlstm$ was trained with Adam with a $\b_1 = 0.9$ and we reduced this to $\b_1 = 0.5$ for $\entropyadam$.}

>>>>

In order to test the efficacy of $\entropysgd$ on recurrent architectures, we train an LSTM network on the Penn Tree Bank (PTB) dataset for word-level text prediction. This dataset contains about one million words divided into a training set of about $930,000$ words, a validation set of $74,000$ words and $82,000$ words with a vocabulary of size $10,000$. As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}. Our network called $\ptblstm$ consists of two layers with $1500$ hidden units each unrolled for $35$ time steps and a dropout of probability $0.65$; this is a large network with about $66$ million weights. It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch. \pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which matches that of the original authors.

@@

Next, we train an LSTM to perform character-level text-prediction. As a dataset, following the experiments of~\citet{karpathy2015visualizing}, we use the text of War and Peace by Leo Tolstoy which contains about $3$ million divided into training ($80\%$), validation ($10\%$) and test ($10\%$) sets. We use an LSTM consisting of two layers of $128$ hidden units unrolled for $50$ time steps and a vocabulary of size $87$. We train the baseline with Adam with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of $1.227 \pm 0.006$.

Let us note that tuning the momentum hyper-parameter was crucial to getting good results on RNNs. While the SGD baseline on $\ptblstm$ does not use momentum and in fact does not train well with momentum, we set it to $0.5$ for $\entropysgd$. On the other hand, the baseline for $\charlstm$ was trained with Adam with a $\b_1 = 0.9$ and we set $\b_1 = 0.5$ for $\entropyadam$. In contrast to this observation on RNNs, our experiments on CNNs above always used a momentum of $0.9$.
